---
title: "07 autocorrelation within reads"
author: "Izaskun Mallona at Robinsonlab"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    code_download: true
    number_sections: true
    df_print: kable
    theme: lumen

# Aim

Postprocesses  `methtutple_bismark_read_correlation.sh`.

```
The trimming data down to 75bp limits the analysis of neighbourhood effects particularly the presence of non-CpG methylation (CpH) in the context of CpG methylation on the same strand. In other words how much is the methylation of CpG’s and neighbouring CpH methylation influenced by flanking sequences on the same molecule.
```

# Set up

Knitted under `R_LIBS=/home/Shared_taupo/Rlib/release-3.11-lib /usr/local/R/R-4.0.0/bin/R`


## Libraries

```{r}
suppressPackageStartupMessages({
    library(pheatmap)
    library(knitr)
    library(ggplot2)
    library(reshape2)
    library(DT)
    library(MASS)
    library(ggExtra)
})
```

## Data load

```{r}
WD <- file.path('/home', 'Shared_s3it', 'imallona', 'cg_context',
                'nar_review', 'autocorrelation')

```

We downsize the data to:
- only `NTUPLES` tuples  (as the CNN context readouts are pretty large).
- We filter tuples with at least two methylation events within the read (to have some data for CNN). That is, those with from columns 9


```{r}

NTUPLES <- 1e6
d <- list()

header <- c('chr', 'strand', 'pos1', 'pos2', 'MM', 'MU', 'UM', 'UU',
            'region1', 'motif1', 'region2', 'motif2')

## 9 is the first tuple column
## column 72 is all unmeth, so shouldn't be included
## 69 is the last column with two Ms at least
## awk '{x=0;for(i=9;i<=69;i++) if($i>=2) x+=1; if(x>1)print}' SRR1274743_1_cutadapt_sickle_bismark_pe.CG_CHG_CHH.6.tsv | less

## awk '{x=0;for(i=9;i<=69;i++) if($i>=1) x+=1; if(x>1) print}' SRR1274743_1_cutadapt_sickle_bismark_pe.CG_CHG_CHH.6.tsv | less

## d$cg <- read.table(pipe(
##     sprintf("awk '{x=0;for(i=9;i<=69;i++) if($i>=1) x+=1; if(x>1) print}' %s",
##             file.path(WD, 'SRR1274743_1_cutadapt_sickle_bismark_pe.CG.6.tsv'))),
##     col.names = header,
##     nrows = NTUPLES)

## d$ch <- read.table(file.path(WD, 'tmp_SRR1274743_1_cutadapt_sickle_bismark.deduplicated.CG_CHG_CHH.2.tsv.gz'),
##                    col.names = header,
##                    sep = '\t',
##                    nrows = NTUPLES)

## d$cg <- read.table(file.path(WD, 'tmp_SRR1274743_1_cutadapt_sickle_bismark_pe.CG.2.tsv.gz'),
##                    col.names = header,
##                    sep = '\t',
##                    nrows = NTUPLES)


d$cnn <- read.table(file.path(WD, 'tmp_SRR1274743_chr10.CG_CHG_CHH.2.tsv.gz'),
                   col.names = header,
                   sep = '\t',
                   nrows = NTUPLES) 
```

# Collapse into frequencies



```{r}
for (item in names(d)) {
    d[[item]]$dist <- d[[item]]$pos2 - d[[item]]$pos1
    d[[item]]$cometh_prop <- d[[item]]$MM/(d[[item]]$UM + d[[item]]$MM + d[[item]]$UU)
}
```

```{r}

lapply(d, function(x) table(x$cometh_prop == 1))

```

```{r}
item <- 'cnn'

boxplot(list(MM = d[[item]][d[[item]]$cometh_prop == 1, 'dist'],
             UU = d[[item]][(d[[item]]$cometh_prop > 0 &
                             d[[item]]$cometh_prop < 1), 'dist'],
             UU = d[[item]][d[[item]]$cometh_prop == 0, 'dist']))
```    


Plot events of at least one methylation event by distance and by motif?


```{r}
for (item in names(d)) {
    d[[item]]$type <- paste(tolower(substr(d[[item]]$motif1, 4, 5)),
                            tolower(substr(d[[item]]$motif2, 4, 5)),
                            sep = '_')

    d[[item]]$type_simple <- 'ch_ch'
    d[[item]]$type_simple <- ifelse(
        test = (tolower(substr(d[[item]]$motif1, 4, 5)) == 'cg' |
                tolower(substr(d[[item]]$motif1, 4, 5)) == 'cg'),
        yes = 'cg_ch',
        no = 'ch_ch')
    
    d[[item]]$type_simple[d[[item]]$type == 'cg_cg'] <- 'cg_cg'

    plot(x = d[[item]][(d[[item]]$MM > 0 | d[[item]]$UM > 0 | d[[item]]$UM > 0), 'dist'],
         d[[item]][(d[[item]]$MM > 0 | d[[item]]$UM > 0 | d[[item]]$UM > 0), 'cometh_prop'],
         pch = 19,
         col = as.numeric(as.factor(
             d[[item]][(d[[item]]$MM > 0 | d[[item]]$UM > 0 | d[[item]]$UM > 0),
                       'type_simple'])))

    ss10 <- smooth.spline(
        d[[item]][(d[[item]]$MM > 0 | d[[item]]$UM > 0 | d[[item]]$UM > 0), 'dist'],
        d[[item]][(d[[item]]$MM > 0 | d[[item]]$UM > 0 | d[[item]]$UM > 0), 'cometh_prop'])
    

    lines(ss10, col = "gray")
}
```

Stratify?

```{r, fig.width = 8, fig.height= 8}
## methylated at least once
m <-  d[[item]][(d[[item]]$MM > 0 | d[[item]]$UM > 0 | d[[item]]$UM > 0),]
ggplot(m, aes(dist, cometh_prop)) + geom_point() + facet_grid(~type_simple) + 
    geom_smooth()
```


```{r}
p <- ggplot(m[m$type_simple == 'cg_cg',], aes(dist, cometh_prop)) +
    geom_point() +
        xlab('') +
        ylab('') +
    theme(legend.position = "none")

ggMarginal(p, margins = 'y',
           groupColour = FALSE, groupFill = FALSE)

```

Now plot according to the nucleotide in the key position.

<!-- ```{r} -->

<!-- library(hexbin) -->
<!-- library(RColorBrewer) -->
<!-- rf <- colorRampPalette(rev(brewer.pal(11,'Spectral'))) -->
<!-- r <- rf(32) -->

<!-- h <- hexbin(d$cg[,c('dist', 'cometh_prop')]) -->
<!-- plot(h, colramp = rf) -->

<!-- ``` -->

<!-- ```{r} -->
<!-- d$cg$logdist <- log(d$cg$dist +1, base = 10) -->
<!-- d$cg$asin_cometh_prop <- asin(d$cg$cometh_prop) -->

<!-- hlog <- hexbin(d$cg[,c('dist', 'asin_cometh_prop')]) -->
<!-- plot(hlog, colramp = rf) -->
<!-- ``` -->



<!-- ```{r} -->
<!-- prop <- list() -->

<!-- for (item in names(d)) { -->
<!--     tuples <- grep('M|U', header, value = TRUE) -->
    
<!--     prop[[item]] <- data.frame(prop = colSums(d[[item]][,tuples]) / sum(d[[item]][,tuples]), -->
<!--                                status = tuples) -->

   
<!-- } -->
<!-- ``` -->


<!-- ```{r} -->

<!-- plot(y = prop$cg$prop, x = 1:nrow(prop$cg), xaxt = 'n', xlab = '', type ='p', col = 'blue', -->
<!--      pch = 0) -->
<!-- lines(y = prop$mixed$prop, x = 1:nrow(prop$cg), type ='p', pch = 4) -->

<!-- axis(1, at = 1:nrow(prop$cg), labels = prop$cg$status, las = 2) -->
<!-- ``` -->


<!-- ```{r} -->

<!-- plot(y = prop$cg$logprop, x = 1:nrow(prop$cg), xaxt = 'n', xlab = '', type ='p', -->
<!--      col = 'blue', pch = 0, ylim = c(min(min(prop$cg$logprop, prop$mixed$logprop)), -->
<!--                                      max(max(prop$cg$logprop, prop$mixed$logprop)))) -->

<!-- lines(prop$mixed$logprop, x = 1:nrow(prop$mixed), type ='p', pch = 4) -->
      
<!-- axis(1, at = 1:nrow(prop$cg), labels = prop$cg$status, las = 2) -->
<!-- ``` -->

<!-- Plot proportion of Ms vs distance between them -->

<!-- first get the tuples that are different from 0 -->
<!-- then get the coordinates -->
<!-- then get the distance between all possible pairs of Ms -->


<!-- ```{r} -->
<!-- dists <- list() -->
<!-- for (item in names(d)) { -->
<!--     dists[[tuple]] <- list() -->
<!--     for (tuple in tuples) { -->
<!--         ## at least two methylation events -->
<!--         if (length(grep('M', strsplit(tuple, '')[[1]])) >= 2) { -->
<!--             ## this gets the relative position of an `M` within a tuple, e.g. UMMMU -> 2,3,4 -->
<!--             ## grep('M', strsplit(tuple, '')[[1]]) -->
<!--             ## this the coordinates of those -->
<!--             ## d[[item]][,paste0('pos', grep('M', strsplit(tuple, '')[[1]]))] -->
<!--             ## and this the distances between the leftmost and all other events -->
            
<!--             dists[[item]][[tuple]] <- apply(d[[item]][,paste0('pos', grep('M', strsplit(tuple, '')[[1]]))], 2, -->
<!--                                             function(x) return(x -d[[item]][,paste0('pos', grep('M', strsplit(tuple, '')[[1]]))][,1])) -->
            
<!--         } -->
<!--     } -->
<!-- } -->

<!-- ``` -->


<!-- ```{r} -->

<!-- item <- 'cg' -->
<!-- curr <- 'MMUUUU' -->

<!-- plot(x= as.vector(dists[[item]][[curr]]), y=prop[[item]][[curr]]$prop) -->

<!-- ``` -->

# Export objects


# Timestamp

```{r sessionInfo, cache = FALSE}
date()
sessionInfo()
devtools::session_info()

```
